### 2021-04-24 latex-teaching/notes.tex
Teaching Notes Template

\documentclass[11pt]{article}

\title{Class Notes}
\author{\normalsize Kyle Butts\\{\footnotesize Univ. of Colorado, Boulder}}
\date{\footnotesize\today}

% Include Theme
\input{preamble.tex}

% Set Theme Colors (for colors look at preamble.tex or define your own)
\colorlet{accent}{indigo}
\colorlet{heading}{indigo!70!black}
\colorlet{emphasis}{indigo}

\begin{document}

\maketitle

\section{Normal Distribution}

This is another paragraph, contains some text to test the paragraph
interlining, paragraph indentation and \itemtag[red]{Marginal Utility} is some other features. Also, is easy to see how new paragraphs are defined by simply entering a 
double blank space.

\itemtag{Marginal Utility} is \outoffive{Rating}{4}

\subsection{Large of Law Numbers}

There are custom boxes for definitions and examples:

\begin{def_box}[accent]
    \textbf{Definition} My definition text here
\end{def_box}

\begin{example_box}[accent]{Bandwidth and noise}
    A band limited signal cannot be time limited and a time-limited signal cannot be band limited and options.
\end{example_box}

These boxes can be put in any color desired:

\begin{def_box}[dark-maroon]
    \textbf{Definition} My definition text here
\end{def_box}

\begin{example_box}[dark-maroon]{Bandwidth and noise}
    A band limited signal cannot be time limited and a time-limited signal cannot be band limited and options.
\end{example_box}

\end{document}