### 2023-03-22 latex-article/article.tex
Update article

\documentclass[12pt]{article}

% Include Theme
\input{preamble.tex}

% \maketitle info
\title{Example Paper}
\author{
    Author One%
    \thanks{School. Email: \href{mailto:email@gmail.com}{email@gmail.com}.} \ and Author Two%
    \thanks{School. Email: \href{mailto:email@gmail.com}{email@gmail.com}.}
}
\date{\today}

% pdf info
\hypersetup{pdftitle={Example Paper}, pdfauthor={Kyle Butts}}

% Conditionally include comments. Switch to \togglefalse
\newtoggle{INCLUDECOMMENTS}
\toggletrue{INCLUDECOMMENTS}
% \togglefalse{INCLUDECOMMENTS}

\newcommand{\kyle}[1]{\iftoggle{INCLUDECOMMENTS}{{\color{cranberry}\textbf{Kyle:} #1}}{}}

\begin{document}

% Title Page -------------------------------------------------------------------
\maketitle
\begin{abstract}

  \par~\par\noindent
  {\color{asher}JEL-Classification:} ...
  \par\noindent
  {\color{asher}Keywords:} ...
  \par\vspace{-2.5mm}
\end{abstract}

\newpage

% Paper ------------------------------------------------------------------------

% ------------------------------------------------------------------------------
\section{Introduction}
% ------------------------------------------------------------------------------

% ------------------------------------------------------------------------------
\newpage~\bibliography{references.bib}
% ------------------------------------------------------------------------------

% ------------------------------------------------------------------------------
\newpage~\appendix
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\renewcommand{\thefigure}{\Alph{section}\arabic{figure}}
\renewcommand{\thetable}{\Alph{section}\arabic{table}}
% ------------------------------------------------------------------------------

\end{document}