### 2023-10-09 referee-response/responses.tex
Response letter to referee reports

\documentclass[12pt]{article}
% Should copy over the .sty files from latex-article
\usepackage{../latex-article/paper}
\usepackage{../latex-article/math}
\usepackage{referee-response}

% Conditionally display thoughts (hide by switching to `\boolfalse`)
\booltrue{INCLUDECOMMENTS}
\newcommand{\kyle}[1]{\coauthorComment[Kyle]{#1}}

% (hide by switching to `\boolfalse`)
\booltrue{INCLUDETODOS}

% Reference main document ------------------------------------------------------
% Note: I compile to `auxiliary/` to avoid cluttering the main directory. You might neeed to delete `auxiliary/`
% \usepackage{xr}
% \externaldocument[main-]{auxiliary/article}

\begin{document}

% Title ------------------------------------------------------------------------
\begin{center}
    {\large\bf Paper Title}

    \emph{Authors' Responses to Referee Comments}

    \emph{Journal} (\texttt{Review Number})
\end{center}
% ------------------------------------------------------------------------------

\noindent Thank you for the additional suggestions. The reviewer's comment is marked with a left border. Our responses follow below each comment.

\section*{Summary of Revisions}

% ------------------------------------------------------------------------------
\NewRef{Referee 1}{R1}
% ------------------------------------------------------------------------------

\begin{refcomment}
  Comment 
\end{refcomment}

\begin{todos}
  Thoughts / plan of action. These can be hidden by toggling \texttt{INCLUDETODOS} to \texttt{\textbackslash boolfalse}. 
\end{todos}

\begin{response}
  Response
\end{response}

\begin{refcomment}
  Comment 
\end{refcomment}

\begin{todos}
  Thoughts / plan of action. 
\end{todos}

\begin{response}
  Response
\end{response}

% ------------------------------------------------------------------------------
\NewRef{Referee 2}{R2}
% ------------------------------------------------------------------------------

\end{document}